Randomized trace estimation for parameter-dependent matrices applied to spectral density approximation

### 2024-08-19 paper/analysis.tex
make chebyshev-nystrom compatible with general parametric trace estimation
add remark for stable-rank bound in Hutchinson

\begin{theorem}{Girard-Hutchinson estimator for parameter-dependent matrices}{hutchinson}
    For a symmetric matrix $\mtx{B}(t) \in \mathbb{R}^{n \times n}$ that depends continuously on $t \in [a, b]$, consider the  Girard-Hutchinson estimator~\refequ{equ:hutchinson-trace-estimator} with an $n\times n_{\mtx{\Psi}}$ Gaussian random matrix $\mtx{\Psi}$. Then for any $p \geq 1, p \in \mathbb{R}$ and $\gamma \geq 1$ the bound 
    \begin{equation} \label{eq:ghbound}
        \int_{a}^{b} |\Trace(\mtx{B}(t)) - \Hutch{\mtx{\Psi}}(\mtx{B}(t))|~\mathrm{d}t \leq \gamma c \left( \sqrt{\frac{p}{n_{\mtx{\Psi}}}} + \frac{p}{n_{\mtx{\Psi}}}  \right) \int_{a}^{b} \lVert \mtx{B}(t) \rVert _F~\mathrm{d}t.
    \end{equation}
    holds with probability at least $1 - \gamma^{-p}$ and for $c = 8 e^{\sfrac{1}{e}} (1 + \sqrt{\pi}/2)$.
    In particular, given $\varepsilon \in (0, 1)$ and $\delta \in (0, e^{-1})$, the bound $\int_{a}^{b} | \Trace(\mtx{B}(t)) - \Hutch{\mtx{\Psi}}(\mtx{B}(t)) | ~\mathrm{d}t \leq \varepsilon \int_{a}^{b} \lVert \mtx{B}(t) \rVert _F~\mathrm{d}t$ holds 
    with probability at least $1-\delta$ if $n_{\mtx{\Psi}} = \mathcal{O}(\varepsilon^{-2} \log(\delta^{-1}))$.
\end{theorem}

@@

>>>>

\begin{theorem}{Girard-Hutchinson estimator for parameter-dependent matrices}{hutchinson}
    For a symmetric matrix $\mtx{B}(t) \in \mathbb{R}^{n \times n}$ that depends continuously on $t \in [a, b]$, consider the  Girard-Hutchinson estimator~\refequ{equ:hutchinson-trace-estimator} with an $n\times n_{\mtx{\Psi}}$ Gaussian random matrix $\mtx{\Psi}$. Then for any $p \geq 1, p \in \mathbb{R}$ and $\gamma \geq 1$ the bound 
    \begin{equation} \label{eq:ghbound}
        \int_{a}^{b} |\Trace(\mtx{B}(t)) - \Hutch{\mtx{\Psi}}(\mtx{B}(t))|~\mathrm{d}t \leq \gamma c \left( \sqrt{\frac{p}{n_{\mtx{\Psi}}}} + \frac{p}{n_{\mtx{\Psi}}}  \right) \int_{a}^{b} \lVert \mtx{B}(t) \rVert _F~\mathrm{d}t
    \end{equation}
    holds with probability at least $1 - \gamma^{-p}$ and for $c = 4 (1 + \sqrt{1 + \pi/4})$.
    In particular, given $\varepsilon \in (0, 1)$ and $\delta \in (0, e^{-1})$, the bound $\int_{a}^{b} | \Trace(\mtx{B}(t)) - \Hutch{\mtx{\Psi}}(\mtx{B}(t)) | ~\mathrm{d}t \leq \varepsilon \int_{a}^{b} \lVert \mtx{B}(t) \rVert _F~\mathrm{d}t$ holds 
    with probability at least $1-\delta$ if $n_{\mtx{\Psi}} = \mathcal{O}(\varepsilon^{-2} \log(\delta^{-1}))$.
\end{theorem}

@@

\begin{remark}
    If $\mtx{A}(t)$ is nonzero and symmetric positive semi-definite for all $t \in [a, b]$, then it follows that with probability at least $1 - \delta$
    \begin{equation}
        \int_{a}^{b} \frac{| \Trace(\mtx{A}(t)) - \Hutch{\mtx{\Psi}}(\mtx{A}(t)) |}{\Trace(\mtx{A}(t))} ~\mathrm{d}t < \varepsilon,
    \end{equation}
    if $n_{\mtx{\Psi}} \geq 2 c^2 e^2 \log(\delta^{-1}) \varepsilon^{-2} (\int_{a}^{b} \mu(t)^{\sfrac{1}{2}}~\mathrm{d}t)^2$ where $\mu(t) = \lVert \mtx{A}(t) \rVert _2 / \Trace(\mtx{A}(t))$. Up to constant factors, this matches the equivalent result for constant matrices from \cite[Remark 2]{cortinovis-2022-randomized-trace}.
\end{remark}
